Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-12-04 papers/integration/omm_integ.tex
Adds work in progress on discussion

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{brain_model} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{ccg_model}(a) \&
(b). Fig.~\ref{brain_model} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{ccg_model}(a).}

@@

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{sbg}. \rvw{In the brainstem model, we use the word `channel'
to mean a set of populations of neurons which are involved in
actuating a single extraocular muscle. SBG channels are arranged in
pairs, actuating opposing muscles. There is one pair of channels which
actuates the superior and inferior rectus muscles, causing vertical
rotations of the eye in a roughly parasaggital plane (the eye moves up
or down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation engines can be
utilised, including DAMSON \citep{richmond_damson_2015}, GeNN
\citep{nowotny_flexible_2011,nowotny_spineml_2014} and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
\citep{cope_spineml_2_brahms_2015-1}. \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a
suitable time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model there is no such
mechanism to reduce activity in SC\_deep, and elsewhere.  Although a
successful, accurate saccade towards a target luminance will remove
the excitation which caused the activity in SC\_deep by bringing the
target luminance within the masked, foveal region, the activity in SC
decays too slowly to avoid additional saccadic movements. We found it
necessary to hypothesise an inhibitory feedback mechanism from the SBG
to the brain model.  This is shown in Fig.~\ref{sbg}, which indicates
how the output from the inhibitory burst neurons (IBN) of the SBG
model are used to feed back an inhibitory signal to the SC\_deep,
thalamus and FEF populations in the brain model, resetting them ready
for the next saccade.
% Might have to justify the above a bit more.

@@

The aim of this study was to demonstrate the importance of modelling
neurological systems \e{in concert with} the biomechanical systems
with which they have evolved in parallel. We hypothesised that by
combining existing neurophysiological models with an accurate model of
a musculo-skeletal system, and then `closing the loop' by allowing the
movements of the virtual muscles to modulate sensory feedback to the
brain model, shortcomings in the constituent models would be revealed,
leading to new knowledge. To demonstrate the validity of this
hypothesis, we built an integrated model and then identified the
modifications which were necessary to give it the ability to make
accurate movements under one type of stimulus. We then examined its
behaviour with other stimuli.

We chose the oculomotor model as a basis for this study because it has
only three degrees of freedom, making it one of the simplest
musculo-skeletal systems. Furthermore, eye movements fall into several
well-defined categories, each being controlled by separate brain
circuits, we were therefore justified in modelling a system which
produced only saccadic eye movements. Nevertheless, we are aware that
we did not create a complete model of the system; no treatment of the
cerebellum was attempted, justified because cerebellum appears to have
only a minor effect on saccade accuracy \citep{dean_adaptive_2008},
probably correcting for slow to medium timescale changes in the
physical dynamics of the eyeball \citep{dean_learning_1994}.

@@

The third issue raised by the integration of the component models of
the saccadic system has, like the return saccades, to do with
resetting activity. In this case, rather than the reset of activity in
the TN population in the brainstem, it is the question of how the
activity in the \e{brain} model should be reset after each
saccade. When a target luminance is projected onto the World
population in the model, this induces activity which `reverberates' in
loops through FEF, basal ganglia, SC and thalamus. The brainstem
contains a mechanism to limit the timescale of a saccade (inhibitory
feedback from EBN, via IBN to LLBN; see Fig.~\ref{sbg}). However, if
the activity in SC is not reset, then following the completion of the
first saccade, a series of subsequent `staircase' saccades will be
executed. There needs to be a mechanism to extinguish activity in SC,
but also in FEF and thalamus, as activity in either of these
populations can build up and eventually cause repeat activity in SC
and another saccade. We added hypothetical inhibitory feedback
connections to our model, such that the IBN populations in the SBG
would inhibit activity in FEF, thalamus and SC\_deep (Fig.~\ref{sbg}),
preventing the occurrence of staircase saccades.

@@

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./figures/Brain_Stem_1channel.png}
\end{center}
\textbf{\refstepcounter{figure}\label{sbg} Figure \arabic{figure}.}
{ One pair of channels of the saccadic burst generator (SBG) for left
(cyan) or right (green) movements. Collicular activity in SC\_avg
excites the channels via SBG weight maps. Each box represents a neural
population and shows the population name, the number of neural
elements (here 2500 or 1) and the SpineML component name; \e{LIN} for
Leaky integrator or \e{integrator}. Key: LLBN: Long lead burst
neurons; IBN: Inhibitory burst neurons; OPN: Omnipause neurons; EBN:
Excitatory burst neurons; TN: Tonic neurons; MN: Motoneurons.}
\end{figure}

>>>>

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{brain_model} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{brain_model} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

@@

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{fig:mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{fig:mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{fig:sbg}. \rvw{In the brainstem model, we use the word `channel'
to mean a set of populations of neurons which are involved in
actuating a single extraocular muscle. SBG channels are arranged in
pairs, actuating opposing muscles. There is one pair of channels which
actuates the superior and inferior rectus muscles, causing vertical
rotations of the eye in a roughly parasaggital plane (the eye moves up
or down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation engines can be
utilised, including DAMSON \citep{richmond_damson_2015}, GeNN
\citep{nowotny_flexible_2011,nowotny_spineml_2014} and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
\citep{cope_spineml_2_brahms_2015}. \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

@@

The saccadic burst generator model was originally conceived with the
assumption of a step input, which returns to zero activity at a
suitable time to curtail the saccade and avoid staircase
saccades \citep{gancarz_neural_1998}. In our model there is no such
mechanism to reduce activity in SC\_deep, and elsewhere.  Although a
successful, accurate saccade towards a target luminance will remove
the excitation which caused the activity in SC\_deep by bringing the
target luminance within the masked, foveal region, the activity in SC
decays too slowly to avoid additional saccadic movements. We found it
necessary to hypothesise an inhibitory feedback mechanism from the SBG
to the brain model.  This is shown in Fig.~\ref{fig:sbg}, which indicates
how the output from the inhibitory burst neurons (IBN) of the SBG
model are used to feed back an inhibitory signal to the SC\_deep,
thalamus and FEF populations in the brain model, resetting them ready
for the next saccade.
% Might have to justify the above a bit more.

@@

The aim of this study was to demonstrate the importance of modelling
neurological systems \e{in concert with} the biomechanical systems
with which they have evolved. We hypothesised that by
combining existing neurophysiological models with an accurate model of
a musculo-skeletal system, and then `closing the loop' by allowing the
movements of the virtual muscles to modulate sensory feedback to the
brain model, shortcomings in the constituent models would be revealed,
leading to new knowledge.
%
\rvw{We call this closed-loop approach \e{computational
neurobehaviour}, meaning a complete, behaving model, with attention
paid to the biological accuracy of each brain and biomechanical
sub-system. There already exist many closed-loop \e{robotic} systems which
receive sensory input from the world, process that input and generate
behaviour by activating motor
systems
\citep{yu_development_2004,fend_artificial_2004,pearson_whiskerbot:_2007}. We
now consider whether robotic systems which model biological
components \e{in hardware} could fall within our new category.
%
Using a number of examples, we will attempt to illustrate what we mean
by computational neurobehaviour. We'll consider which examples fall
into the new category and which are covered by other fields of
robotics or computational neuroscience. We will then attempt to
provide a definition of computational neurobehaviour.}

\rvw{\cite{pearson_whiskerbot:_2007} describes a wheeled robot which has a
biomimetic whisker sensory system, along with a biomimetic neural
system imitating the behaviour of the rat's sensory processing systems
and controlling the movement of the robot. \cite{fend_artificial_2004}
is a similar, wheeled, whiskered robot, with a repertoire of three
behaviours organised in a subsumption architecture. In both robots,
actions that are selected within the brain model drive a
non-biologically accurate wheel motor control algorithm to achieve
rotational and translational movements. Although both have
sensory and processing systems which are guided by biology, the
non-biological motor control stage prevents us from considering these
as being studies of computational neurobehaviour.}
%
\rvw{\cite{yu_development_2004} report on a biomimetic fish robot,
whose motor system closely resembles that of the real fish. The robot
is able to operate in a closed-loop mode, where sensory input is
provided to the non-biomimetic control algorithms from overhead
cameras, but its control system is also able to operate in open-loop
mode. Neither the control system, nor the sensory system are
biophysically accurate and we would not describe this study as
computational neurobehaviour. Nevertheless, the biophysically accurate
motor system could certainly form part of a computational
neurobehavioural study of swimming behaviour in the fish, if it were
combined with suitable sensory input and sensory processing models.}
%
\rvw{\cite{knips_neural_2017} is a report of a reach-and-grasp robot
arm controlled via a dynamic neural field brain model. The sensory
input for this system---its `eyes'---is a Microsoft Kinect sensor; it
also has somatosensory feedback from the fingers of the robot's
hand. The neural field `brain' controls the seven degrees of freedom
of the arm to carry out the reach-and-grasp action. While this robot
has closed-loop control and is clearly inspired by biology, it remains
a study of robotics and of the improvement of the control of the
robot's reach-and-grasp function, rather than a study which aims to
learn more about the biology of a primate arm. For this reason, we
would describe the study of \cite{knips_neural_2017} as embodied
robotics or neuro-robotics.}
%
\rvw{\cite{arena_motor-skill_2017} describes a
robotic insect system based on the fly species \e{Drosophila
Melanogaster}. It has biomimetic insect legs implemented in a virtual
robot and a neuro-mimetic brain. Closing the loop is a visual sensory
input system which is able to determine the distance between the robot
and an obstacle and the obstacle's height. The visual system is not
described in detail and may not be biologically modelled (much in the
way that the visual processing stream is not modelled in detail in the
present work). Nevertheless, this study may reasonably be described as
compuational neurobehaviour; the authors demonstrate that the virtual
robot is able to learn to climb obstacles in a realistic manner and
suggest that the model may be compared with experimental data from
future Drosophila experiments addressing obstacle climbing and
learning.}

\rvw{To summarise, we give a definition of computational
neurobehaviour as \e{the study of biological sensory-motor system
behaviour using biophysically accurate models of sensory input, brain
and motor sub-systems operating in a closed-loop}. The sub-system
models may be implemented in software or hardware, provided that they
attempt to reproduce and explain features of the biological
sub-system.}

To demonstrate the validity of the computational neurobehaviour
approach, we built an integrated model and then identified the
modifications which were necessary to give it the ability to make
accurate movements under one type of stimulus. We then examined its
behaviour with other stimuli. We chose the oculomotor model as a basis
for this study because it has only three degrees of freedom, making it
one of the simplest musculo-skeletal systems. Furthermore, eye
movements fall into several well-defined categories, each being
controlled by separate brain circuits, we were therefore justified in
modelling a system which produced only saccadic eye
movements. Nevertheless, we are aware that we did not create a
complete model of the system; no treatment of the cerebellum was
attempted, justified because cerebellum appears to have only a minor
effect on saccade accuracy \citep{dean_adaptive_2008}, probably
correcting for slow to medium timescale changes in the physical
dynamics of the eyeball \citep{dean_learning_1994}.

@@

The third issue raised by the integration of the component models of
the saccadic system has, like the return saccades, to do with
resetting activity. In this case, rather than the reset of activity in
the TN population in the brainstem, it is the question of how the
activity in the \e{brain} model should be reset after each
saccade. When a target luminance is projected onto the World
population in the model, this induces activity which `reverberates' in
loops through FEF, basal ganglia, SC and thalamus. The brainstem
contains a mechanism to limit the timescale of a saccade (inhibitory
feedback from EBN, via IBN to LLBN; see Fig.~\ref{fig:sbg}). However, if
the activity in SC is not reset, then following the completion of the
first saccade, a series of subsequent `staircase' saccades will be
executed. There needs to be a mechanism to extinguish activity in SC,
but also in FEF and thalamus, as activity in either of these
populations can build up and eventually cause repeat activity in SC
and another saccade. We added hypothetical inhibitory feedback
connections to our model, such that the IBN populations in the SBG
would inhibit activity in FEF, thalamus and SC\_deep (Fig.~\ref{fig:sbg}),
preventing the occurrence of staircase saccades.

@@

% Reviewer 1's point 2.
\rvw{Address reviewer's question of ``whether inaccuracies in the
biomechanical model may introduce
errors that propagate through the closed loop system and cannot be
counteracted by the nervous system part of the simulation. How can you
(or someone using this or related models) exclude introducing
(biomechanical) properties that are not present in the biological
system, or make sure to not miss critical biomechanical properties
that might help remedy some of the shortcomings of the closed loop
model?''.}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.6\textwidth]{./figures/Brain_Stem_1channel.png}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:sbg} Figure \arabic{figure}.}
{ One pair of channels of the saccadic burst generator (SBG) for left
(cyan) or right (green) movements. Collicular activity in SC\_avg
excites the channels via SBG weight maps. Each box represents a neural
population and shows the population name, the number of neural
elements (here 2500 or 1) and the SpineML component name; \e{LIN} for
Leaky integrator or \e{integrator}. Key: LLBN: Long lead burst
neurons; IBN: Inhibitory burst neurons; OPN: Omnipause neurons; EBN:
Excitatory burst neurons; TN: Tonic neurons; MN: Motoneurons.}
\end{figure}
